---
title: "Beta Binomial Model - Your Use Case"
format: html
editor: visual
---

# Welcome to Today's Lecture

```{r}
library(tidyverse)
library(bayesrules)
```

We have done a deep-dive into the [Beta Binomial model for Michellele's election support](https://www.bayesrulesbook.com/chapter-3) with the shiny app in the `beta_binomial_election_app.R` script. Now, I would like you and a teammate to create a use-case that is appropriate for the usage of this model.

## Example

This example will help guide you in your use case, I will leave a skeleton outline of the bottom of this qmd for you to fill out.

### Prior

Every four years, Americans go to the polls to cast their vote for President of the United States. Consider the following scenario. “Michelle” has decided to run for president and you’re her campaign manager for the state of Minnesota. As such, you’ve conducted 30 different polls throughout the election season. Though Michelle’s support has hovered around 45%, she polled at around 35% in the dreariest days and around 55% in the best days on the campaign trail

π = Michelle's true election support

```{r}
plot_beta(45, 55)
```

Statement = "Before any new poll, we believe there's a 95% chance Michelle's true support π is between 35% and 55%, with 45% as our single best guess."

Prior effective sample size = 100 pseudo voters

### Likelihood

**New data arrives:** We commission a new poll of *n* Minnesotans and record *y* , the number who support Michelle. We poll 50 voters in which 30 (60%) supported her -- noticeably higher than her 45% prior average.

```{r}
plot_binomial_likelihood(y = 30, n = 50)
```

Likelihood Question: "For each possible value of π — Michelle's true underlying support — how compatible is that value with having produced exactly 30 out of 50 in this poll?"

Likelihood Statement: "Taken on its own, this poll makes values of π near 60% far more plausible than values near 45% — the data alone points to a higher number than we'd been expecting"

### Posterior

```{r}
plot_beta_binomial(alpha = 45, beta = 55, y = 30, n = 50)
```

Statement = "There's a 95% probability that Michelle's true support π lies between 42.0% and 58.0%."

Weight = prior (worth \~100 pseudo-voters) contributes 67% of the weight, and our new poll (n = 50 voters) contributes the other 33%.

## Your Turn

With your teammate, invent a use-case appropriate for the Beta Binomial Model. You do not need to find a dataset, just follow the kind of thinking from the text and example. Repeat the steps above describing in context your prior, likelihood, and posterior with statements and plots. Tips:

- Come up with 3 -5 ideas on your own

- Sketch them out and think critically in how they differ from the example but may still fit the Beta Binomial Modeling architecture

- Pitch your sketches to a free AI Chat of your choosing (I use Claude) and have it critique your use case to see if the Beta Binomial Model is appropriate and accurately addresses your question

- Note the assumptions/limitations of your model based on AI feedback

- **Remember**: "[All models are wrong, but some are useful](https://en.wikipedia.org/wiki/All_models_are_wrong)"

Things to keep on mind:

- Is π a single, clearly defined quantity — not "kind of about deposits, kind of about attendance"?

- Does your data collection method actually measure that population, or a biased subset of it (who responds, who's easy to ask)?

- Is π genuinely uncertain — or is it something you could already know exactly (a completed census), making the model unnecessary?

- Is your prior based on real past counts (pseudo-counts) or an elicited belief — and can you say why you chose that one?

### Prior

Set the scene...

the main question we are going to answer is what percent of NBA players played on division 1 rosters before going pro. the prior is the roster from 1980 and the new data is the roster from 2026. we saw that less players in the 2026 roster played on division 1 rosters than in 1980.

π = The proportion of NBA players that played in NCAA D1 and then ended up playing professional.

```{r}
plot_beta(alpha = 234 , beta = 30 )
```

Statement = "Before any new data we believe there's a 95% chance π is between 85% and 92%, with 89% as our single best guess."

Prior effective sample size = 264

### Likelihood

**New data arrives:** We commission new data of 540 and record 340, the number of collegiate who ended up on the 2026 roster. We gather data points in which 340 (62%) played in the NBA in 2026 -- noticeably lower than the 89% prior average.

```{r}
plot_binomial_likelihood(y = 340, n = 540)
```

Likelihood Question: "For each possible value of π —Division 1 athletes playing in the NBA — how compatible is that value with having produced exactly 340 out of 540 in this data collection?"

Likelihood Statement: "Taken on its own, this data makes values of π near 61% far more plausible than values near 89% — the data alone points to a lower number than we'd been expecting"

### Posterior

```{r}
plot_beta_binomial(alpha = 234, beta = 30, y = 340, n = 540 )
```

Statement = "There's a 95% probability that π lies between 68.2% and 74.5%."

Weight = prior (worth 264 ) contributes 33% of the weight, and our new data (n = 540) contributes the other 67%.
